### 2020-04-22 final_report/tbr_final_report-metadata.tex
add bitmap ukaea logo placeholder, enable draft mode

%-------------------------------------------------------------------------
% This file contains the title, author and abstract.
% It also contains all relevant document numbers if needed.
%-------------------------------------------------------------------------

% Partner Logo
% put the name of the logo image file found in graphics path
% TODO: make this a vector logo, wait for reply from Vignesh
\DISPartnerLogo{ukaea_logo}

% Title
\DISTitle{Surrogate Modelling of\\the Tritium Breeding Ratio}

% Draft version:
% If given, adds draft version on front page, a 'DRAFT' box on top of each other page, 
% and line numbers.
% TODO: Comment or remove in final version.
\DISVersion{0}

% Abstract - % directly after { is important for correct indentation
\DISAbstract{%
  The abstract of my report.
  % TODO: write this bit
}

% Authors and list of contributors to the analysis
\usepackage{authblk}
\author[a]{Petr Mánek}
\author[a]{Graham Van Goffrier}
\affil[a]{University College London}

% DIS reference code if ever used
%\DISRefCode{UCLCDTDIS-2019-XX}