### 2020-04-22 final_report/tbr_final_report-metadata.tex

%-------------------------------------------------------------------------
% This file contains the title, author and abstract.
% It also contains all relevant document numbers if needed.
%-------------------------------------------------------------------------

% Partner Logo
% put the name of the logo image file found in graphics path
% TODO: make this a vector logo, wait for reply from Vignesh
\DISPartnerLogo{ukaea_logo}

% Title
\DISTitle{Surrogate Modelling of\\the Tritium Breeding Ratio}

% Draft version:
% If given, adds draft version on front page, a 'DRAFT' box on top of each other page, 
% and line numbers.
% TODO: Comment or remove in final version.
\DISVersion{0}

% Abstract - % directly after { is important for correct indentation
\DISAbstract{%
  Surrogate modelling techniques have been employed in diverse areas of research to resolve the limitations of resource-expensive procedure with significantly cheaper approximations. In this internship project, we sought to develop a fast and high-quality surrogate model for simulations of the tritium breeding ratio (TBR) for under-design Tokamak nuclear fusion reactors. We investigated possibilities for dimensional reduction on the parameter space of a Monte Carlo TBR model in use at the UK Atomic Energy Authority. We investigated possibilities for dimensional reduction on the parameter space of this model, and reviewed $N$ classes of surrogate models for potential applicability. Here we present the performance and scaling properties of these surrogate models, the best of which demonstrated an accuracy of $X$ and a mean prediction time of $X$, representing a relative speedup $X$ with respect to the MC model. We further present a novel adaptive sampling algorithm, QASS, capable of interfacing with any of the individual studied models. Our preliminary testing on a toy TBR theory has demonstrated the efficacy of this algorithm for speeding up the surrogate modelling process.
}

% Authors and list of contributors to the analysis
\usepackage{authblk}
\author[a]{Petr Mánek}
\author[a]{Graham Van Goffrier}
\affil[a]{University College London}

% DIS reference code if ever used
%\DISRefCode{UCLCDTDIS-2019-XX}