### 2020-04-23 final_report/appendix_online.tex
add outline of appendix B

TODO: point reader to various online repositories, briefly describe them

\begin{enumerate}
	\item data repository
	\item sampling repository
	\item regression repository
	\item hyperopt repository
	\item docs repository
\end{enumerate}