### 2020-04-30 final_report/tbr_final_report-metadata.tex

%-------------------------------------------------------------------------
% This file contains the title, author and abstract.
% It also contains all relevant document numbers if needed.
%-------------------------------------------------------------------------

% Partner Logo
% put the name of the logo image file found in graphics path
\DISPartnerLogo{ukaea_logo}

% Title
\DISTitle{Surrogate Modelling of\\the Tritium Breeding Ratio}

% Draft version:
% If given, adds draft version on front page, a 'DRAFT' box on top of each other page, 
% and line numbers.
%\DISVersion{0}

% Abstract - % directly after { is important for correct indentation
\DISAbstract{%
 The tritium breeding ratio (TBR) is an essential quantity for the design of
 modern and next-generation Tokamak nuclear fusion reactors. Representing the
 ratio between tritium fuel generated in breeding blankets and fuel consumed
 during reactor runtime, the TBR depends in a
 complex manner on reactor geometry and material properties. We explored the
 training of surrogate models to produce a cheap but high-quality approximation
 for a Monte Carlo TBR model in use at the UK Atomic Energy Authority. We
 investigated possibilities for dimensional reduction on the feature space of
 this model, reviewed 9~families of surrogate models for potential
 applicability and performed hyperparameter optimisation. Here we present the
 performance and scaling properties of these
 models, the best of which demonstrated~$R^2=\num{0.998}$ and a mean
 prediction time of~$\SI{0.898}{\micro\second}$, representing a relative speedup~$\num{8659251} \times$
 with respect to the expensive MC model. We further present a novel adaptive
 sampling algorithm, QASS, capable
 of interfacing with any of the individual studied models. Our preliminary
 testing on a toy TBR theory has demonstrated the efficacy of this algorithm for
 accelerating the surrogate modelling process.
}

% Authors and list of contributors to the analysis
\usepackage{authblk}
\author[a]{Petr Mánek}
\author[a]{Graham Van Goffrier}
\affil[a]{University College London}

% DIS reference code if ever used
%\DISRefCode{UCLCDTDIS-2019-XX}