### 2020-05-21 paper/ucl_tbr_paper.tex
add provisional title, institution addresses

\documentclass[12pt]{iopart}

\usepackage{iopams}
\usepackage[colorinlistoftodos]{todonotes} % TODO: remove me before submitting

\providecommand{\keywords}[1]{\textbf{\textit{Index terms---}} #1}

\begin{document}
\title[Fast regression of the Tritium Breeding Ratio in Tokamaks]{Fast regression of the
Tritium Breeding Ratio in Tokamak fusion reactors} % TODO: revise this

\author{G~Van Goffrier$^1$ and P~Mánek$^{1,2}$, V~Gopakumar$^3$, N~Nikolau$^1$, J~Shimwell$^3$, I~Waldmann$^1$}

\address{$^1$ Department of Physics and Astronomy, University College London, Gower Street, London WC1E~6BT, UK}
\address{$^2$ Institute of Experimental and Applied Physics, Czech Technical University, Husova 240/5, Prague 110~00, Czech Republic}
\address{$^3$ UK Atomic Energy Authority, Culham Science Centre, OX14~3DB Abingdon, UK}
\ead{williams@ucl.ac.uk}

\begin{abstract}
% TODO: revise this
... 
\end{abstract}

\keywords{magnetic moment, solar neutrinos, astrophysics}
\submitto{\jpg}
\maketitle

\todo{This is a comment that will appear in the margin}

\end{document}