### 2020-06-19 talk_ucl_cdt_200622/conclusion.tex
add outline for CDT talk on Monday

\begin{frame}
	\frametitle{Conclusion}
	\begin{itemize}
		\item % TODO
	\end{itemize}
\end{frame}