### 2020-06-19 talk_ucl_cdt_200622/ucl_tbr_talk.tex
set graphics path, include UKAEA logo on the title slide

\documentclass{beamer}

% Main theme configuration
\usetheme{ucl}
\useoutertheme[small]{ucltitlebanner}
\setbeamercolor{banner}{bg=darkblue}

% Packages and settings
\graphicspath{{./fig_final_report/}{./fig_paper/}{./fig/}}

% Set a smaller indent for description environment
\setbeamersize{description width=2em}

% Remove nav symbols (and shift any logo down to corner)
\setbeamertemplate{navigation symbols}{\vspace{-2ex}}

% Metadata
\titlegraphic{\includegraphics[width=0.12\paperwidth]{ukaea_logo}}
\title[Surrogate Modelling of the Tritium Breeding Ratio]{Surrogate Modelling of the Tritium Breeding Ratio}
\author[Petr Mánek \and Graham Van Goffrier]{Petr Mánek \and Graham Van Goffrier}
\institute[UCL]{%
	Centre for Doctoral Training in Data Intensive Science \\ %
	University College London
}
\date{22nd June 2020}

\begin{document}

% Title slide
\begin{frame}
  \titlepage%
\end{frame}

% Body of the presentation is organized in separate files
\section{Introduction}
\input{introduction}

\section{Decoupled Approach}
\input{decoupled}

\section{Adaptive Approach}
\input{adaptive}

\section{Conclusion}
\input{conclusion}

\end{document}