### 2020-06-21 talk_ucl_cdt_200622/conclusion.tex

\begin{frame}
	\frametitle{Conclusion}

	Decoupled approach:

	\begin{itemize}
		\item
			Tuned and compared surrogates from 9~state-of-the-art families.
		\item
			Found heuristic: GBTs for $<10^4$~points and
			ANNs for $\geq10^5$~points.
		\item
			Fastest found surrogate predicts TBR with standard deviation of
			error~$\num{0.033}$ in~\SI{0.898}{\micro\second}, which is~$8\cdot
			10^6\times$~faster than Paramak.
		\item
			While this used 500K~datapoints, we found surrogates with
			comparable properties with as little as 10K~datapoints.
	\end{itemize}

	\vspace{0.5em}

	Adaptive approach:
	\begin{itemize}
		\item 
			TODO % TODO Graham
		\item
			TODO % TODO Graham
		\item
			TODO % TODO Graham
		\item
			TODO % TODO Graham
	\end{itemize}

	\vspace{0.5em}

	Presented methods portable $\rightarrow$ can be used as cheap
	approximation of any simulation or black box function.
\end{frame}