### 2020-06-21 talk_ucl_cdt_200622/ucl_tbr_talk.tex

\documentclass{beamer}

% Main theme configuration
\usetheme{ucl}
\useoutertheme[small]{ucltitlebanner}
\setbeamercolor{banner}{bg=darkblue}

% Packages and settings
\usepackage{siunitx}
\usepackage{isotope}
\usepackage{hyperref}
\graphicspath{{./fig_final_report/}{./fig_paper/}{./fig/}}

\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}

% Set a smaller indent for description environment
\setbeamersize{description width=2em}

% Remove nav symbols (and shift any logo down to corner)
\setbeamertemplate{navigation symbols}{\vspace{-2ex}}

% Title slide logos
\newlength{\titlegraphicheight}
\setlength{\titlegraphicheight}{0.15\paperheight}
\titlegraphic{%
	\includegraphics[height=\titlegraphicheight]{cdt_dis_logo}%
	\hspace{15pt}%
	\includegraphics[height=\titlegraphicheight]{ukaea_logo}%
}

% Use title slide at section beginning
\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

% Metadata
\title[Surrogate Modelling of the Tritium Breeding Ratio]{Surrogate Modelling of the Tritium Breeding Ratio}
\author[Petr Mánek \and Graham Van Goffrier]{Petr Mánek \and Graham Van Goffrier}
\institute[UCL]{%
	Centre for Doctoral Training in Data Intensive Science \\ %
	University College London
}
\date{22nd June 2020}

\begin{document}

% Title slide
\begin{frame}
  \titlepage%
\end{frame}

% Body of the presentation is organized in separate files
\input{introduction}

\section{Decoupled Approach}
\input{decoupled}

\section{Adaptive Approach}
\input{adaptive}

\input{conclusion}

\begin{frame}
	\vspace{2em}

	{\centering
		\begin{beamercolorbox}[sep=8pt,center]{title}
			\usebeamerfont{title} Thank you for listening! \par%
		\end{beamercolorbox}
	}

	\vspace{2em}

	Petr Mánek \dotfill \email{petr.manek.19@ucl.ac.uk}

	Graham Van~Goffrier \dotfill \email{graham.vangoffrier.19@ucl.ac.uk}

	\vspace{2em}

	{\footnotesize
	Further reading:
	\begin{itemize}
		\item
			\href{https://github.com/ucl-tbr-group-project/documentation/raw/pinboards-submitted/elevator_pitch/elevator_pitch.pdf}{Single page abstract (available online)}.
		\item
			\href{https://github.com/ucl-tbr-group-project/documentation/raw/pinboards-submitted/paper/ucl_tbr_paper.pdf}{Journal article, currently in internal pre-submission review (available online):\\
			\textit{Fast Regression of the Tritium Breeding Ratio in Fusion
		Reactors}}.
		\item
			\href{https://github.com/ucl-tbr-group-project/documentation/raw/published/final_report/tbr_final_report.pdf}{Industry group project final report (available online)}.
		\item
			All models, plots, training data, source code and technical documentation.
			\url{https://github.com/ucl-tbr-group-project}
	\end{itemize}
	}

\end{frame}

\end{document}