### 2020-10-07 talk_ucl_cdt_201008/conclusion.tex
Small updates to adaptive sampling wording

\begin{frame}
	\frametitle{Conclusion}

	\begin{block}{Decoupled Approach}
		\begin{itemize}
			\setlength\itemsep{0em}
			\item
				Heuristic: \alert{GBTs} for $<10^4$~samples,
				\alert{ANNs} for $\geq10^5$~samples.
			\item
				Fastest found surrogate evaluates TBR in~\SI{0.898}{\micro\second}
				with error~$\num{0.033}$. This is roughly~\alert{$8\cdot
				10^6\times$~faster} than Paramak.
			\item
				Found surrogates with comparable properties with
				\alert{$\approx$10K~samples}.
		\end{itemize}
	\end{block}

	\begin{block}{Adaptive Approach}
		\begin{itemize}
			\setlength\itemsep{0em}
			\item
				New theoretical approach \alert{QASS} developed, based on MCMC.
			\item 
				\alert{$60\%$ decrease} in evaluation MAE demonstrated.
			\item
				\alert{$6\%$ decrease} in expensive TBR samples needed.
		\end{itemize}
	\end{block}

	\begin{itemize}
		\item 
			\alert{Portable} methods $\rightarrow$ cheap approximation of any simulation.
		\item
			Article in IOP \alert{Journal of Nuclear Fusion} (pending).
		\item
			Included as a benchmark in the \alert{SciML Collaboration}.
	\end{itemize}

\end{frame}