### 2022-08-08 paper/acknowledgements.tex
Removed textcolor and minor grammar adjustments.

PM \& GVG were supported by the STFC UCL Centre for Doctoral Training in Data Intensive
Science (grant no. ST/P006736/1). GVG was funded by the UCL Graduate Research and Overseas Research Scholarships.

This project was supported by the EU Horizon 2020 research \& innovation
programme [grant No 758892, ExoAI]. N.~Nikolaou acknowledges the support of the NVIDIA Corporation’s GPU grant.

This work has been carried out within the framework of the EUROfusion consortium and has received funding from the Euratom research and training programme 2014-2018 and 2019-2020 under grant agreement No 633053. The views and opinions expressed herein do not necessarily reflect those of the European Commission.

This work has been partly funded by the Institutional support for the
development of a research organization (DKRVO, Czech Republic).

This work has also been part-funded by the RCUK Energy Programme (grant number EP/I501045/1).